An Abstraction-based Approach to Network Synthesis

### 2016-10-31 paper/abstraction/paper.tex
start with compilation example
remove conceret methane

\sysname builds on the Propane network configuration synthesis language to let
operators specify \emph{network-wide} routing policy at a high level of abstraction.
A \sysname policy consists of a sequence of ranked constraints on the paths traffic should 
or should not take for certain IP prefix destinations.

\subsection{Concrete Methane}

Consider again the datacenter from figure~\ref{fig:example}. To configure the datacenter routing
policy, we would start by writing the following definition in \sysname:
%
\begin{code}
\Define Routing =
    \{P1 \Path \End(T1)
     P2 \Path \End(T2)
     P3 \Path \End(T3)
     P4 \Path \End(T4)
     P5 \Path \End(T5)
     P6 \Path \End(T6)
     P7 \Path \End(T7)
     P8 \Path \End(T8)
     \True \Path \Exit(Peer1 \Prefer Peer2) \}
\end{code}
\noindent%

This defines a group of constraints for basic routing behavior. The first line states 
that traffic destined for prefix $P1$ must follow a path that ends up at destination router $T1$.
A similar constraint is added for each of the prefixes P[2-8]. The final rule matches all other
prefix destinations and allows traffic to follow a path that leaves the datacenter through either
of its peers (Peer1 or Peer2) with a preference for leaving through Peer1. The \Prefer symbol indicates
that traffic should satisfy the constraint on the left whenever possible, and only resort to the backup (right)
when this is not possible due to failures.

Next we can add the locality constraint for prefixes P[5-8]. We do this by adding another group of constraints:
%
\begin{code}
\Define Locality =
    \{P5 | P6 | P7 | P8 \Path \Always(\In)\}
\end{code}
\noindent%

The Locality constraint applies to any traffic destined for a local prefix and adds
the constraint that the traffic must follow a path that matches an internal location 
(\IE, a router inside the datacenter) at each hop of the path. This prevents traffic 
from ever entering from or leaving through an external peer.

Finally, for all traffic in the datacenter, we will disallow valley routing. This constraint
can be captured with the following additions to the policy:
%
\begin{code}
\Define Tor = \{T1,T2,T3,T4,T5,T6,T7,T8\}
\Define Agg = \{A1,A2,A3,A4,A5,A6,A7,A8\}
\Define Spn = \{S1,S2\}
\Define ValleyFree =
    \{\True \Path \Novalley(Tor,Agg,Spn)\}
\end{code}
\noindent%

The first three lines define groups of routers. The top of rack (Tor) routers T[1-8], the 
aggregation routers (Agg) A[1-8], and the spine routers (Spn) S[1-2]. The next line 
adds a new constraint that applies to all traffic and ensures that traffic never follows a 
path that forms a valley between any of these three groups (\EG, going from Spn to Agg to Spn again).
As we will see in section \ryan{reference}, each of these constraints such as the valley free constraint
are really just syntactic sugar for regular expressions over topology locations.

The final policy is declared to as the conjunction of each of these constraints:
%
\begin{code}
\Define Main = 
    Routing & 
    Locality & 
    ValleyFree & 
    \Agg(PG, \In -> \Out)
\end{code}
\noindent%
The last part of the policy declares that we want to perform aggregation for global prefixes
by summarizing them into the more general prefix PG at the border of the datacenter
(\IE along any edge starting inside our datacenter and ending outside).

From this policy, \sysname can synthesize device-level BGP configurations for every router in 
the network that guarantee correct forwarding behavior under all possible failure combinations. 
Synthesizing configurations for each device from a network-wide specification both simplifies 
network configuration by writing policy in one place rather than dozens and simultaneously prevents 
a large class of errors associated with manual configuration by construction. For example, 
the compiler ensures configurations are always in sync and prevents the operator from having 
to manually consider failure scenarios.

@@

Although the concrete \sysname policy can help prevent a number of configuration errors, 
the policy is intricately linked to the actual structure of the concrete network.
For example, each prefix in the routing constraint is enumerated with its concrete destination 
location, and the definition of the valleyfree constraint depends on the actual routers that 
comprise each tier of the datacenter. 
%
This is unfortunate since any expansion of our datacenter network will require a corresponding 
change to the policy. Furthermore, there is no guarantee that a safe policy can be found 
for the expanded network, and even if one can be found, it may require changes to every
configuration in the network -- a huge barrier to adoption in practice.

Abstract \sysname solves this problem in two ways. First, it allows operators to abstract
the network into a role-based topology and write a routing specification over this abstraction instead.
Second, it introduces prefix template variables into the language to allow writing policies that are 
parameterized over prefix destination.

\begin{figure}[t!]
  \centering
  \includegraphics[width=.6\columnwidth]{figures/example3}
  \caption{Role-based abstraction for the datacenter in Figure~\ref{fig:example1}.}
  \label{fig:example3}
  \vspace{-1em}
\end{figure}

@@

%This  statement defines a set of \emph{ranked paths}, which includes
%all paths (and only those paths) for which traffic exits our network
%through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.
%The paths that exit through \CD{R1}
%are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that
%leave through \CD{Peer} and then \CD{Prov}.  As we
%describe in the next section, the \Exit\ expression, as well as other
%path expressions used later in this section, is simply a shorthand for
%a particular regular expression over paths that is expressible in our policy
%language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among %individual routers. For example, the above constraint could have been written equivalently as
%\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})%
%To associate ranked paths with
%one or more prefixes, we define a \sysname \emph{policy}.
%Within a policy, statements with the form $t\;$\Path$\;p$
%associate the prefixes defined by the predicate $t$ with the set of
%ranked paths defined by the path expression $p$.  In general,
%prefix predicates can be defined by arbitrary boolean combinations
%(and, or, not) of concrete prefixes and community tags.  Here,
%we assume we have already defined the predicate \CD{PCust} for the
%customer prefixes.
%
%In the following code, ranked paths are associated with
%customer prefixes,
%and all other prefixes (\CD{true}). Policy statements are processed in
%order with earlier policy statements taking precedence over later
%policy statements. Hence, when the predicate \True{} follows
%the statement involving \CD{PCust}, it is interpreted as
%\CD{\True{} \AND{} \NOT{}PCust}.%

%\begin{code}
%\Define Routing =
%    \{PCust \Path Prefs \AND \End(Cust)
%     \True  \Path Prefs \}
%\end{code}
%\noindent%
%Line 2 of this policy
%restricts traffic destined to known customer prefixes (\CD{PCust}) to only follow paths that end at the customer. In addition, it enforces the network-wide preference that traffic leaves through \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}.
%%
%Line 3 applies to any other traffic not matching \CD{PCust} and allows the traffic to leave through any direct neighbor with the usual preference of \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}. To summarize our progress, the \CD{Routing} policy
%implements P1, P3, and P4 from Figure~\ref{fig:example1}.%%

%Since, routing allows transit traffic by default (\EG, traffic entering from
%\CD{Peer} and leaving through \CD{Prov}), we separately define a
%policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
%disjunction (\OR) and negation (\NOT) of constraints.
%First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
%\begin{code}
%\Define transit(X,Y) = \Enter(X{\OR}Y) \AND \Exit(X{\OR}Y)
%\Define cust-transit(X,Y) = \Eventually(X) \AND \Eventually(Y)
%\end{code}
%\noindent
%Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then also leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.%
%\begin{code}
%\Define NoTrans =
%  \{\True \Path \NOT{}transit(Peer,Prov) \AND
%           \NOT{}cust-transit(Cust,Prov)\}
%\end{code}
%\noindent
%The \CD{NoTrans} constraint requires that all traffic not follow a path that transits our network between \CD{Peer} and \CD{Prov}.
%Additionally, it prevents traffic from ever following paths that leave our network and later go through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
%and \CD{NoTrans} simultaneously, we simply conjoin them: \CD{Routing \AND{} NoTrans}.%
%Collectively, the constraints above capture the entire policy. From them, our compiler will generate per-device import and export filters, local preferences,
%MED attributes, and community tags to ensure that the policy is
%implemented correctly under all failures.%%

%Our datacenter example network has three main concerns:
%(1) traffic for the prefix allocated to each top-of-rack router must be able to reach that router,
%(2) local services must not leak outside the datacenter, and
%(3) aggregation must be performed on global prefixes to reduce churn
%in the network.%
%\sysname allows modular specification of each of these constraints. The first constraint is about prefix ownership---we want traffic only for certain prefixes to end up at a particular location. The following definition captures this intent.%%

%\begin{code}
%\Define Ownership =
%    \{PG1 \Path \End(A)
%     PG2 \Path \End(B)
%     PL1 \Path \End(E)
%     PL2 \Path \End(F)
%     \True \Path \End(\Out)\}
%\end{code}
%\noindent
%This definition says that traffic for prefix \CD{PG1} is allowed to follow only paths that
%end at router \CD{A}; traffic for \CD{PG2},
%but not \CD{PG1}, must
%end at router \CD{B}; and so on.
%Any traffic destined for a prefix that is not a part of the datacenter should be allowed to leave the datacenter and end at some external location, which is otherwise unconstrained.
%The special keyword \Out{} matches any location outside the datacenter network, while the keyword \In{} will match any location inside the network.%
%For the second constraint, we define another policy:%
%\begin{code}
%\Define Locality =
%    \{PL1 \OR PL2 \Path \textbf{only}(\In)\}
%\end{code}
%\noindent
%%
%This definition says that traffic for local
%prefixes only follows paths that are internal to the network at each hop.
%This constraint guarantees that the services remain accessible only to locations
%inside the datacenter.%
%As in the backbone example, we can logically conjoin these constraints
%to specify the network-wide policy.
%However, in addition to constraints on the shape of paths,
%\sysname allows the operator to specify constraints on the BGP control plane itself.
%For instance, a constraint on aggregation is included to ensure that
%aggregation for global prefixes is performed from
%locations inside (\In) the network to locations outside (\Out).
%In this case, \CD{PG1} and \CD{PG2} will use the aggregate \CD{PG}
%(which we assume is defined earlier)
%when advertised outside the datacenter.%
%\begin{code}
%Ownership \AND{} Locality \AND{} \Agg(PG, \In \Link \Out)
%\end{code}%
%Once \sysname compiles the policy, it is guaranteed to remain compliant under all possible failure scenarios, modulo any aggregation-induced black holes. In the presence of aggregation, the \sysname compiler will also efficiently find a lower bound on the number of failures required to create an aggregation-induced black hole.

@@

@@

%\begin{figure*}[t!]
%  \begin{minipage}[t]{.5\linewidth}%

%  \hdr{Topology}{}
%  \vspace*{-2\baselineskip}%

%  \includegraphics[width=.6\columnwidth]{figures/topology}%

%  \hdr{Policy Automata}{}
%  \vspace*{-0.5\baselineskip}%

%  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.4cm,minimum size=1cm]
%    \state{0}{$0$}{              }{}
%    \state{1}{$1$}{right of=0}{}
%    \state{2}{$2$}{right of=1}{}
%    \state{3}{$3$}{right of=2}{}
%    \state{4}{$4$}{right of=3}{}
%    \state{5}{$5$}{right of=4}{accepting}
%    \transition{0}{1}{out}{}
%    \transition{1}{2}{D}{}
%    \transition{2}{3}{C}{}
%    \transition{3}{4}{A}{}
%    \transition{4}{5}{W}{}
%  \end{tikzpicture}%

%  \vspace*{0.5em}%
%

%  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
%    \state{0}{$0$}{              }{}
%    \state{1}{$1$}{right of=0}{}
%    \state{2}{$2$}{right of=1}{}
%    \state{3}{$3$}{right of=2}{}
%    \state{4}{$4$}{right of=3}{accepting}
%    \transition{0}{1}{out}{}
%    \transition{1}{2}{in}{}
%    \transition{2}{2}{A,C,D,E}{loop above}
%    \transition{2}{3}{B}{}
%    \transition{3}{3}{B}{loop above}
%    \transition{3}{2}{A,C,D,E}{bend left}
%    \transition{3}{4}{W}{}
%  \end{tikzpicture}
%  \end{minipage}
%  %
%  ~~
%  ~~
%  %
%  \begin{minipage}[t]{.5\linewidth}
%  \hdr{Product Graph IR}{}
%  \vspace*{-1\baselineskip}
%  %
%  \includegraphics[width=.8\columnwidth]{figures/productgraph}
%  \end{minipage}%

%  \hrulefill
%  \vspace*{.4em}%

%  \caption{Product graph construction for policy $(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$.}
%  \label{fig:example-compilation}
%  %\vspace{-1em}
%\end{figure*}

@@

Figure~ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract compilation. Notably, the abstract analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to Z3 remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by Z3 to minimize the same constraints with different parameters.

>>>>

As we have seen, writing correct network configurations is challenging task.
Operators must manually decompose network-wide policies into device configurations,
reason about all combination of network failures, parameterize configurations to 
allow for change, and write configurations in a way amenable to incremental change.
We now give an overview of \sysname, a system that facilitates all of these tasks.

%\subsection{Concrete Methane}%

%Consider again the datacenter from figure~\ref{fig:example}. To configure the datacenter routing
%policy, we would start by writing the following definition in \sysname:
%%
%\begin{code}
%\Define Routing =
%    \{P1 \Path \End(T1)
%     P2 \Path \End(T2)
%     P3 \Path \End(T3)
%     P4 \Path \End(T4)
%     P5 \Path \End(T5)
%     P6 \Path \End(T6)
%     P7 \Path \End(T7)
%     P8 \Path \End(T8)
%     \True \Path \Exit(Peer1 \Prefer Peer2) \}
%\end{code}
%\noindent%%

%This defines a group of constraints for basic routing behavior. The first line states 
%that traffic destined for prefix $P1$ must follow a path that ends up at destination router $T1$.
%A similar constraint is added for each of the prefixes P[2-8]. The final rule matches all other
%prefix destinations and allows traffic to follow a path that leaves the datacenter through either
%of its peers (Peer1 or Peer2) with a preference for leaving through Peer1. The \Prefer symbol indicates
%that traffic should satisfy the constraint on the left whenever possible, and only resort to the backup (right)
%when this is not possible due to failures.%

%Next we can add the locality constraint for prefixes P[5-8]. We do this by adding another group of constraints:
%%
%\begin{code}
%\Define Locality =
%    \{P5 | P6 | P7 | P8 \Path \Always(\In)\}
%\end{code}
%\noindent%%

%The Locality constraint applies to any traffic destined for a local prefix and adds
%the constraint that the traffic must follow a path that matches an internal location 
%(\IE, a router inside the datacenter) at each hop of the path. This prevents traffic 
%from ever entering from or leaving through an external peer.%

%Finally, for all traffic in the datacenter, we will disallow valley routing. This constraint
%can be captured with the following additions to the policy:
%%
%\begin{code}
%\Define Tor = \{T1,T2,T3,T4,T5,T6,T7,T8\}
%\Define Agg = \{A1,A2,A3,A4,A5,A6,A7,A8\}
%\Define Spn = \{S1,S2\}
%\Define ValleyFree =
%    \{\True \Path \Novalley(Tor,Agg,Spn)\}
%\end{code}
%\noindent%%

%The first three lines define groups of routers. The top of rack (Tor) routers T[1-8], the 
%aggregation routers (Agg) A[1-8], and the spine routers (Spn) S[1-2]. The next line 
%adds a new constraint that applies to all traffic and ensures that traffic never follows a 
%path that forms a valley between any of these three groups (\EG, going from Spn to Agg to Spn again).
%As we will see in section \ryan{reference}, each of these constraints such as the valley free constraint
%are really just syntactic sugar for regular expressions over topology locations.%

%The final policy is declared to as the conjunction of each of these constraints:
%%
%\begin{code}
%\Define Main = 
%    Routing & 
%    Locality & 
%    ValleyFree & 
%    \Agg(PG, \In -> \Out)
%\end{code}
%\noindent%
%The last part of the policy declares that we want to perform aggregation for global prefixes
%by summarizing them into the more general prefix PG at the border of the datacenter
%(\IE along any edge starting inside our datacenter and ending outside).%

%From this policy, \sysname can synthesize device-level BGP configurations for every router in 
%the network that guarantee correct forwarding behavior under all possible failure combinations. 
%Synthesizing configurations for each device from a network-wide specification both simplifies 
%network configuration by writing policy in one place rather than dozens and simultaneously prevents 
%a large class of errors associated with manual configuration by construction. For example, 
%the compiler ensures configurations are always in sync and prevents the operator from having 
%to manually consider failure scenarios.

@@

\sysname builds on the Propane network configuration synthesis language, 
which lets operators specify a \emph{network-wide} routing policy, by adding
support for compiling policies parameterized over different topology.
In particular, a \sysname policy consists of predicates matching
certain classes of traffic, paired with a list of ranked constraints 
on the paths traffic should prefer to take for the predicate.

Consider again the datacenter from earlier. Ideally, we would like
to find device configuration templates for each router that, taken together, ensure correct forwarding
behavior for both of the datacenter topologies in figures~\ref{fig:example} and~\ref{fig:example2}.

\sysname achieves this by allowing the operator to abstract the network into 
a role-based topology that captures a class of concrete topologies and then write a 
routing specification over this abstraction. Figure~\ref{fig:example3} shows an example of one 
such abstraction for the datacenter networks. The network is partitioned into 5 different roles:
local tor (TL), global tor (TG), local aggregator (AL), global aggregator (AG), and spine (S).
Each router in the concrete topology maps to exactly one abstract role.

%Although the concrete \sysname policy can help prevent a number of configuration errors, 
%the policy is intricately linked to the actual structure of the concrete network.
%For example, each prefix in the routing constraint is enumerated with its concrete destination 
%location, and the definition of the valleyfree constraint depends on the actual routers that 
%comprise each tier of the datacenter. 
%
%This is unfortunate since any expansion of our datacenter network will require a corresponding 
%change to the policy. Furthermore, there is no guarantee that a safe policy can be found 
%for the expanded network, and even if one can be found, it may require changes to every
%configuration in the network -- a huge barrier to adoption in practice.

%Abstract \sysname solves this problem in two ways. First, it allows operators to abstract
%the network into a role-based topology and write a routing specification over this abstraction instead.
%Second, it introduces prefix template variables into the language to allow writing policies that are 
%parameterized over prefix destination.

\begin{figure}[t!]
  \centering
  \includegraphics[width=.6\columnwidth]{figures/example3}
  \caption{Role-based abstraction for the datacenter in Figure~\ref{fig:example}.}
  \label{fig:example3}
  \vspace{-1em}
\end{figure}

@@

\subsection{Language}

\ryan{TODO: introduce language + semantics}

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{.2\linewidth}
  \hdr{Policy Automata}{}
  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{accepting}
    \transition{0}{0}{out}{loop above}
    \transition{0}{1}{Peer1}{}
    \transition{1}{1}{in}{loop above}
  \end{tikzpicture}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2cm,minimum size=.5cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{accepting}
    \transition{0}{0}{out}{loop above}
    \transition{0}{1}{Peer2}{}
    \transition{1}{1}{in}{loop above}
  \end{tikzpicture}%
  \end{minipage}
  %
  ~~~~~~
  %
  \begin{minipage}[t]{.45\linewidth}
  \hdr{Concrete Topology}{}
  \includegraphics[width=\columnwidth]{figures/example}
  \end{minipage}%
  %
  \begin{minipage}[t]{.35\linewidth}
  \hdr{Abstract Topology}{}
  \includegraphics[width=.75\columnwidth]{figures/example3}
  \end{minipage}%

  \begin{minipage}[t]{.6\linewidth}
  \hdr{Concrete Product Graph}{}
  %\includegraphics[width=.65\columnwidth]{figures/example3}
  \end{minipage}%
  \begin{minipage}[t]{.4\linewidth}
  \hdr{Abstract Product Graph}{}

    \begin{tikzpicture}
      \draw [] (0, 0) circle [radius=0.45] node {start};
      \draw [] (-1, -1) circle [radius=0.45] node {$(S,0,0)$};
      %\draw [] (0, .45) -- (0, 2.05);
    \end{tikzpicture}

  \hrulefill
  \vspace*{.4em}%

  \caption{Product graph construction for policy $true \Path \Exit(\text{Peer1} \Prefer \text{Peer2})$.}
  \label{fig:example-compilation}
  %\vspace{-1em}
\end{figure*}

@@

@@

Figure~\ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract compilation. Notably, the abstract analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to Z3 remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by Z3 to minimize the same constraints with different parameters.
